### 2019-02-21 fantasy_baseball/draftResearch_pt1.Rmd
Init commit

---
title: "Draft Research pt1"
author: "Tom Kain"
date: "February 20, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Info for getting depth chart projections:
[Batters](https://www.fangraphs.com/projections.aspx?pos=all&stats=bat&type=fangraphsdc&team=0&lg=all&players=0)

[Pitchers](https://www.fangraphs.com/projections.aspx?pos=all&stats=pit&type=fangraphsdc&team=0&lg=all&players=0)

Download to CSV via:
`javascript:__doPostBack('ProjectionBoard1$cmdCSV','')`

Downloading manually for now...

```{r}
pitchers <- read_csv("data/depthChartProjections_pitchers_20190220.csv")

batters <- read_csv("data/depthChartProjections_batters_20190220.csv")

pitchers
batters
```