### 2019-03-05 draftResearch_pt1.Rmd
Working on compiling up eligible positions into a comma separated string or vector

---
title: "Draft Research pt1"
author: "Tom Kain"
date: "February 20, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("R/yahoo_api.R")
library(tidyverse)
```

# Position Eligibility
```{r}
(get_players(50) -> tst_results)

tst_results %>%
  select(starts_with("eligible_positions")) %>%
  transpose() %>%
  lapply(as.character) %>%
  lapply(function(x) x[!is.na(x)]) -> pos_complete

# add list rows representing new column as new column on prior df?
tst_results %>%
  add_column(pos_complete) %>%
  select(1, pos_complete)

```