### 2019-03-09 draftResearch_pt2.Rmd
Import fangraphs depth chart projection data

---
title: "Draft Research pt2"
author: "Tom Kain"
date: "March 8, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Custom player rankings

Goals:
1. Import projection data, as automated as possible
2. Generate ranks based on league stat categories using projection data via z-scores controlling for PA/IP
3. Determine a replacement-level performance and scale rankings accordingly

# Import
Info for getting depth chart projections:
[Batters](https://www.fangraphs.com/projections.aspx?pos=all&stats=bat&type=fangraphsdc&team=0&lg=all&players=0)

[Pitchers](https://www.fangraphs.com/projections.aspx?pos=all&stats=pit&type=fangraphsdc&team=0&lg=all&players=0)

Download to CSV via:
`javascript:__doPostBack('ProjectionBoard1$cmdCSV','')`

Downloading manually for now. The scoring categories for our league:
Batters: R, H, HR, RBI, SB, BB, K, OBP, OPS - Not included in data set: E
Pitchers: SV, K, HLD, ERA, WHIP - Not included in data set: NH, QS, BSV, CG, SHO

```{r}
pitchers_file <- "data/depthChartProjections_pitchers_20190308.csv"
batters_file <- "data/depthChartProjections_batters_20190308.csv"

pitchers <- read_csv(pitchers_file) %>%
  select(playerid, 
         Name, 
         Team,
         ADP,
         IP,
         SV:ERA,
         SO,
         WHIP)

batters <- read_csv(batters_file) %>%
  select(playerid,
         Name,
         Team,
         ADP,
         PA,
         H,
         HR:SO,
         SB,
         OBP,
         OPS)
```