### 2019-03-12 draftResearch_pt2.Rmd
Completed adjusted validation

---
title: "Draft Research pt2"
author: "Tom Kain"
date: "March 8, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Custom player rankings

Goals:
1. Import projection data, as automated as possible
2. Generate ranks based on league stat categories using projection data via z-scores controlling for PA/IP
3. Determine a replacement-level performance and scale rankings accordingly

# Import
Info for getting depth chart projections:
[Batters](https://www.fangraphs.com/projections.aspx?pos=all&stats=bat&type=fangraphsdc&team=0&lg=all&players=0)

[Pitchers](https://www.fangraphs.com/projections.aspx?pos=all&stats=pit&type=fangraphsdc&team=0&lg=all&players=0)

Download to CSV via:
`javascript:__doPostBack('ProjectionBoard1$cmdCSV','')`

Downloading manually for now. The scoring categories for our league:
Batters: R, H, HR, RBI, SB, BB, K, OBP, OPS - Not included in data set: E
Pitchers: SV, K, HLD, ERA, WHIP - Not included in data set: NH, QS, BSV, CG, SHO

```{r}
pitchers_raw <- "data/depthChartProjections_pitchers_20190308.csv" %>%
  read_csv()
batters_raw <- "data/depthChartProjections_batters_20190308.csv" %>%
  read_csv()

pitchers <- pitchers_raw %>%
  select(playerid, 
         Name, 
         Team,
         ADP,
         IP,
         SV:ERA,
         SO,
         WHIP)

batters <- batters_raw %>%
  select(playerid,
         Name,
         Team,
         ADP,
         PA,
         H,
         HR:SO,
         SB,
         OBP,
         OPS)
```

# Generate rankings
```{r}
z_score_count <- function(df, stat, invert=FALSE, multiplier=1) {
  stat <- enquo(stat)
  col_nm <- paste0("z_", quo_name(stat))

  df <- df %>% 
    mutate(s_z = (!! stat - mean(!! stat)) / sd(!! stat))

  if (invert) {
    df %>%
      mutate(s_z = -s_z) -> df
  }

  if (multiplier != 1) {
    df %>%
      mutate(s_z = multiplier * s_z) -> df
  }

  df %>%
   rename(!! col_nm := s_z)
}

z_score_rate <- function(df, stat, quantity, invert=FALSE, multiplier = 1) {
  stat <- enquo(stat)
  quantity <- enquo(quantity)
  col_nm <- paste0("z_",quo_name(stat))

  df %>%
    mutate(s_q = (!! stat - mean(!! stat)) * !! quantity) %>%
    mutate(s_z = (s_q - mean(s_q)) / sd(s_q)) -> df

  if (invert) {
    df %>%
      mutate(s_z = -s_z) -> df
  }

  if (multiplier != 1) {
    df %>%
      mutate(s_z = multiplier * s_z) -> df
  }

  df %>%
    rename(!! col_nm := s_z) %>%
    select(-s_q)
}

(z_pitchers <- pitchers %>%
  filter(ADP < 999) %>%
  z_score_count(SV) %>%
  z_score_count(HLD) %>%
  z_score_rate(ERA, IP, TRUE) %>%
  z_score_count(SO) %>%
  z_score_rate(WHIP, IP, TRUE) %>%
  mutate(
    z_sum = pmap_dbl(
      select(., starts_with("z_")), 
      sum
    )
  ) %>%
  arrange(desc(z_sum))
)

(z_batters <- batters %>%
  filter(ADP < 999) %>%
  z_score_count(H) %>%
  z_score_count(HR) %>%
  z_score_count(R) %>%
  z_score_count(RBI) %>%
  z_score_count(BB) %>%
  z_score_count(SO, TRUE) %>%
  z_score_count(SB) %>%
  z_score_rate(OBP, PA) %>%
  z_score_rate(OPS, PA) %>%
  mutate(
    z_sum = pmap_dbl(
      select(., starts_with("z_")), 
      sum
    )
  ) %>%
  arrange(desc(z_sum))
)

z_batters %>% saveRDS("data/z_batters.Rds")
z_pitchers %>% saveRDS("data/z_pitchers.Rds")

z_compare <- z_batters %>% 
  mutate(type = "Batter") %>%
  bind_rows(z_pitchers) %>%
  select(playerid, Name, type, ADP, z_sum) %>%
  mutate(type = if_else(is.na(type), "Pitcher", type)) %>%
  arrange(z_sum %>% desc())

z_compare %>%
  mutate(ovrl_rank = row_number()) %>%
  ggplot(aes(x=ADP,y=ovrl_rank,color=type)) +
  geom_point(aes(text=Name)) +
  geom_abline(slope=1) -> p

plotly::ggplotly(p)

```

### Missing Projection Data
Looking for some proxies for the following: NH, QS, BSV, CG, SHO, E. NH and SHO scarcity probably isn't worth diving in, probably just weight CG higher if I can project that. As for the remainders:

- Errors : research the correlation with Depth Chart's "FLD" metric to see if that's useful enough to weigh lightly. 
- BSV : Look at prior years' data to see the blown save rate and see if there's a predictable relationship to anything
- QS : once we've got a tally (find a source or summarize game logs), same idea as above
- CG : same

##### Errors
```{r}
fielding_standard <- "data/2018_Fielding-Standard.csv" %>%
  read_csv() %>%
  mutate(playerid = playerid %>% as.character())

fielding_advanced <- "data/2018_Fielding-Advanced.csv" %>%
  read_csv() %>%
  mutate(playerid = playerid %>% as.character()) %>%
  select(-(Name:Inn))

fielding <- fielding_standard %>%
  full_join(fielding_advanced, "playerid")

fielding %>%
  select(playerid, Pos, E, Def) %>%
  full_join(batters_raw, "playerid") %>%
  filter(ADP < 999) %>%
  select(Name, Team, Pos, E, Def, Fld) %>%
  filter(!is.na(E)) %>%
  ggplot(aes(x=E,y=Fld)) +
  geom_point()

```
Incomplete listing, probably due to minimum qualified innings, but even with this sample (~ a third of the batters in the depth chart projections), not seeing a pattern. Since there is going to be at least one pitching category I won't be able to project, not having Errors should be ok.

##### CG
```{r}
hist_pitching <- "data/2016-2018_Pitching.csv" %>%
  read_csv()

hist_pitching %>%
  ggplot(aes(x=CG/3)) + 
  geom_histogram(binwidth=1)

hist_pitching %>%
  ggplot(aes(x=CG/3,y=IP/3)) +
  geom_point()
```
Looks like there were only three pitchers in the past three years that averaged 1 CG/year with less than 150 IP/year. Takeaway is to weigh IP a bit more.

##### BSV
```{r}
hist_pitching %>%
  select(Name, Team, G, IP, ERA, SV, BS) %>%
  filter(SV > 0 | BS > 0) -> sv_pitching

sv_pitching %>%
  ggplot(aes(x=SV/3)) +
  geom_histogram(binwidth=1)

sv_pitching %>%
  ggplot(aes(x=SV/3,y=BS/3)) +
  geom_point()

sv_pitching %>%
  ggplot(aes(x=SV/3,y=IP/3)) +
  geom_point()

sv_pitching %>%
  ggplot(aes(x=BS/3,y=IP/3)) +
  geom_point()
```
As expected, not much with the existing data to correlate Blown Saves, so I the takeaway will probably be to more highly weigh ERA and WHIP.

##### QS
Struggling to find meaningful QS data to work with, but again, I'm thinking the takeaway will be more weighting on ERA & WHIP.

##### Z-Score Adjustments
I originally added a multiplier of 3 to two of the pitching categories, so that it would mean there were "nine" categories for both pitching and batting that make up the z_sum. 
The result was that pitching z_sums had a much wider range, to the point where three pitchers had a higher z_sum than Mike Trout, who was coming out roughly as valuable as Justin Verlander. 
Weighting the categories so that there would be an "equal" number of categories felt right, but I wonder if that's giving too much credit to scarce categories like CG/SHO. 
For now, I've set the multiplier to 2. The results are still a bit "scandalous" (Mookie Betts going after Sale/deGrom/Scherzer) but this is a better starting point for a conversation.
Need to take a deeper look into how I'm comparing the relative value between pitcher and catcher rather than setting the multiplier to get results I'm expecting. 
```{r}
(z_pitchers_adj <- pitchers %>%
  filter(ADP < 999) %>%
  z_score_count(SV) %>%
  z_score_count(HLD) %>%
  z_score_rate(ERA, IP, TRUE, multiplier=2) %>%
  z_score_count(SO) %>%
  z_score_rate(WHIP, IP, TRUE, multiplier=2) %>%
  mutate(
    z_sum = pmap_dbl(
      select(., starts_with("z_")), 
      sum
    )
  ) %>%
  arrange(desc(z_sum))
)

(z_compare_adj <- z_batters %>% 
  mutate(type = "Batter") %>%
  bind_rows(z_pitchers_adj) %>%
  mutate(type = if_else(is.na(type), "Pitcher", type)) %>%
  arrange(z_sum %>% desc()) %>%
  mutate(ovrl_rank = row_number()))

z_compare_adj %>% saveRDS("data/z_compare_adj.Rds")

z_compare_adj %>%  
  filter(Name == "Miguel Cabrera" | Name == "Kyle Freeland") %>% 
  View()

z_compare_adj %>%
  ggplot(aes(x=ADP,y=ovrl_rank,color=type)) +
  geom_point(aes(text=Name)) +
  geom_abline(slope=1)-> p_adj

plotly::ggplotly(p_adj)
```

# Validation
```{r}
z_compare_adj %>%
  filter(type != "Batter") %>% 
  select(playerid, Name, Team, ADP, ovrl_rank, z_sum, z_ERA, z_WHIP, z_SO, z_SV, z_HLD, IP, ERA, WHIP, SO, SV, HLD) %>%
  arrange(ADP) %>%
  View()

z_compare_adj %>%
  ggplot(aes(z_sum, color = type)) +
  geom_freqpoly(binwidth=1)

z_compare_adj %>%
  filter(Name %in% c("Blake Snell", "Andrew Benintendi")) %>%
  View()

z_compare_adj %>%
  filter(Name %in% c("James Paxton", "Eugenio Suarez")) %>% 
  View()
```

I don't have any objective measure to compare the value between similarly ranked batters and pitchers. I compared a few pairs of players against other expert's rankings and these largely seem to be in line. I'm at a disadvantage without being able to project Quality Starts. However, since I'm walking in with the likely goal of spending early draft capital on hitting and planning to play the waiver wire for two-start/matchup starters, these results should be sufficient. 
### To Do
1. Develop tiers of players (linear model? nearest neighbor?)
2. Replacement level stuff (new section)
3. Take averages of multiple projection sources? 
4. Build team stats based on selections, available players per round?