### 2019-03-12 draftResearch_pt3.Rmd
Starting join

---
title: "Draft Research pt3"
author: "Tom Kain"
date: "March 11, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Join Data
Let's see if we can join the eligible position data (Yahoo) with the projection data (Fangraphs).

```{r}
yahoo <- readRDS('data/players_position_eligibility.Rds') %>%
  select(name.full, display_position, positions)

fangraphs <- readRDS('data/z_compare_adj.Rds')

df_raw <- fangraphs %>%
  full_join(yahoo, by = c("Name" = "name.full"))

```

```{r}
df_raw %>%
  filter(is.na(display_position)) %>%
  select(Name, positions, ADP, ovrl_rank, z_sum) %>%
  head(25)

df_raw %>%
  filter(Name == "Jose Ramirez") %>%
  select(Name, positions, ADP, ovrl_rank, z_sum) %>%
  head()

fangraphs %>%
  count()
```

### To Do
1. Missing joins in yahoo data (i.e. Jose Ramirez) due to inconsistent accent marks on letters (and maybe other causes too?)
2. Develop tiers of players (linear model? nearest neighbor?)
3. Replacement level stuff (new section)
4. Take averages of multiple projection sources? 
5. Build team stats based on selections, available players per round?