### 2019-03-22 draftPrep.Rmd
Work in progress switching to latest projections

---
title: "2019 Draft Prep"
author: "Tom Kain"
date: "March 19, 2019"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(stop_on_error = 2L)
source("R/yahoo_api.R")
options(tidyverse.quiet=TRUE)
library(tidyverse)
```

# Intro
Rather than loosely apply fantasy baseball expert rankings for traditionally-scored roto leagues, I wanted to take a stab at developing my own rankings based on several factors:

1. league-custom categories used for H2H scoring
2. quality projection data with realistic expectations for variance
3. my preferences for draft strategy and team management throughout the year

To that end, the below document is a walk-through of the various research completed in other R Markdown files found in this repo.

### Scoring Categories
Here are the statistical categories which are used in our H2H categories, 12 team fantasy baseball league. Both team and category record are tracked throughout the year.

##### Batters
- Runs (R)
- Hits (H)
- Home runs (HR)
- Runs Batted In (RBI)
- Stolen Bases (SB)
- Walks (BB)
- Strikeouts (SO/K)
- On-Base Percentage (OBP)
- On-Base Plus Slugging (OPS)
- \*Errors (E)

##### Pitchers
- Saves (SV)
- Strikeouts (SO/K)
- Holds (HLD)
- Earned Run Average (ERA)
- Walks+Hits per Inning (WHIP)
- \*No Hitters (NH)
- \*Quality Starts (QS)
- \*Blown Saves (BS)
- \*Complete Games (CG)
- \*Shut outs (SHO)

*Categories marked with an asterisk (\*) are not inluded in projection data.*

### Projection Data
Based on prior personal success and their construction is the average of two separate projections, I've downloaded the latest Fangraphs Depth Chart preseason projections. This data is joined with Yahoo position eligibility data, since this is our league's host.

### Personal Strategy
After playing fantasy baseball for a few seasons and reading up from various experts (CBS Sports, The Athletic, etc), here are a few takeaways and preferences I prefer to take to heart at draft time as well as during the season.

1. Batter projections tend to be more accurate than pitchers
2. I tend to ride the waiver wire for starting pitching to take advantage of two-start weeks in an effort to find quality starts
3. Since more batting categories are projected than pitching categories, I tend to put more emphasis on getting hitters early in drafts

# Data Set
```{r importFangraphs, include=FALSE}
fangraphs_pitchers <- "data/fangraphs_pitchers.Rds"
fangraphs_batters <- "data/fangraphs_batters.Rds"
pitchers_file <- "data/depthChartProjections_pitchers_20190321.csv"
batters_file <- "data/depthChartProjections_batters_20190321.csv"

if (!file.exists(pitchers_file) || !file.exists(batters_file)) {
  stop("Couldn't find fangraphs export data in the data folder")
}

pitchers <- 
  pitchers_file %>%
  read_csv() %>%
  unique() %>% 
  mutate(type = "pitcher")

batters <- batters_file %>%
  read_csv() %>%
  unique() %>%
  mutate(type = "batter")

fangraphs <- pitchers %>% 
  bind_rows(batters)

fangraphs$Name[fangraphs$Name == 'Enrique Hernandez'] <- "Kike Hernandez"
fangraphs$Name[fangraphs$Name == 'Cedric Mullins II'] <- "Cedric Mullins"
fangraphs$Name[fangraphs$Name == 'Michael Soroka'] <- "Mike Soroka"
fangraphs$Name[fangraphs$Name == 'Nathaniel Lowe'] <- "Nate Lowe"
fangraphs$Name[fangraphs$Name == 'Nathan Karns'] <- "Nate Karns"
fangraphs$Name[fangraphs$Name == 'Zach Britton'] <- "Zack Britton"
fangraphs$Name[fangraphs$Name == 'Joshua James'] <- "Josh James"
fangraphs$Name[fangraphs$Name == 'D.J. Stewart'] <- "DJ Stewart"

batters <- fangraphs %>%
  filter(type=="batter") %>%
  select(playerid,
         Name,
         Team,
         ADP,
         PA,
         H,
         HR:SO,
         SB,
         OBP,
         OPS)

pitchers <- fangraphs %>%
  filter(type=="pitcher") %>%
  select(playerid,
         Name,
         Team,
         ADP,
         IP,
         SV:ERA,
         SO,
         WHIP)

batters %>% saveRDS(fangraphs_batters)
pitchers %>% saveRDS(fangraphs_pitchers)

suppressWarnings(
  rm(fangraphs_pitchers, fangraphs_batters, pitchers_file, batters_file) 
)
```

```{r importYahoo, include=FALSE}
yahoo_file <- "data/players_position_eligibility.Rds"
if (file.exists(yahoo_file)) {
  yahoo_players <- readRDS(yahoo_file)
} else {
  get_players(2500) -> yahoo_players

  position_columns <- yahoo_players %>%
    select(starts_with("eligible_positions")) %>%
    colnames()

  yahoo_players %>%
    mutate(positions = paste(!!! syms(position_columns), sep = "," )) %>%
    mutate(positions = gsub('NA', '', positions)) %>%
    mutate(positions = gsub("^,*|(?<=,),|,*$", "", positions, perl=T)) %>%
    select(-(starts_with("eligible_positions"))) %>%
  select(name.full, display_position, positions) %>%
  mutate(name.full = chartr("áéóíñú", "aeoinu", name.full))  -> yahoo_players

  yahoo_players$name.full[yahoo_players$name.full == 'Daniel Vogelbach'] <- "Dan Vogelbach"
  yahoo_players$name.full[yahoo_players$name.full == 'Shohei Ohtani (Batter)'] <- "Shohei Ohtani"
  yahoo_players$name.full[yahoo_players$name.full == 'Shin-soo Choo'] <- "Shin-Soo Choo"
  yahoo_players$name.full[yahoo_players$name.full == 'Seunghwan Oh'] <- "Seung Hwan Oh"
  yahoo_players$name.full[yahoo_players$name.full == 'JT Riddle'] <- "J.T. Riddle"
  yahoo_players$name.full[yahoo_players$name.full == 'AJ Reed'] <- "A.J. Reed"

  yahoo_players %>% 
    saveRDS(yahoo_file)
}
rm(yahoo_file)
```

```{r}
players_master <- fangraphs %>%
  full_join(yahoo_players, by = c("Name" = "name.full")) %>%
  filter(!is.na(ADP))

players_master %>% 
  mutate(orphan = is.na(display_position) & is.na(positions)) %>%
  filter(orphan) -> orphans

if (!is.na(orphans) && nrow(orphans) > 0) {
  stop(paste0("Join resulted in ",nrow(orphans)," orphaned players"))
}

orphans %>% select(Name)
fangraphs %>% filter(grepl("Genesis",Name))
yahoo_players %>% filter(grepl("Cabrera",name.full))
```