### 2019-04-06 tradeReview.Rmd
Starting to get team roster from Yahoo Fantasy Sports Api

---
title: "Trade Reviews"
author: "Tom Kain"
date: "April 5, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
```

Various trade scenarios to review during the year.

## Early April
Targeting a 3B, considering Chris's team as a trade partner since he has multiple high value 3B and needs 1B, RP. I'll need to

1. Get the latest ROS projection data, calculate z_scores
2. Get the players for both teams with their eligible positions
3. Get draft results
4. Join roster, projection, and draft result data, rescale z_scores appropriately
5. Identify some equitable trade scenarios keeping as close to a net even z_score exchange

```{r}
source("R/yahoo_api.R")
league_string <- paste0("388.l.",Sys.getenv("LEAGUE_KEY"))
get_team_roster(league_string, 8) %>% 
  clean_names() %>% 
  mutate(owner = 'Tom') -> my_team
get_team_roster(league_string, 3) %>% 
  clean_names() %>% 
  mutate(owner = 'Chris') -> chris_team

my_team %>% 
  bind_rows(chris_team) %>% 
  select(owner, name_full, display_position) -> roster_players

```