### 2019-04-06 tradeReview.Rmd
Joined various data sources for first trade review

---
title: "Trade Reviews"
author: "Tom Kain"
date: "April 5, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
```

Various trade scenarios to review during the year.

## Early April
Targeting a 3B, considering Chris's team as a trade partner since he has multiple high value 3B and needs 1B, RP. I'll need to...

1. Get the latest ROS projection data, calculate z_scores
2. Get the players for both teams with their eligible positions
3. Get draft results
4. Join roster, projection, and draft result data, rescale z_scores appropriately
5. Identify some equitable trade scenarios keeping as close to a net even z_score exchange

#### Latest ROS Projection data, calculate z_scores
```{r}
source("R/z_scores.R")
proj_batter_raw <- read_csv("data/depthChartProjections_ros_batters_20190405.csv") 
proj_pitcher_raw <- read_csv("data/depthChartProjections_ros_pitchers_20190405.csv")

proj_batter_raw %>%
  z_score_count(R) %>%
  z_score_count(H) %>%
  z_score_count(HR) %>%
  z_score_count(RBI) %>%
  z_score_count(SB) %>%
  z_score_count(BB) %>%
  z_score_count(SO, TRUE) %>%
  z_score_rate(OBP, PA) %>%
  z_score_rate(OPS, PA) %>%
  mutate(
    z_sum = pmap_dbl(
      select(., starts_with("z_")), 
      sum
    ),
    type = "batter") %>% 
  select(Name, type, z_sum) -> proj_batter

proj_pitcher_raw %>%
  z_score_count(SV) %>%
  z_score_count(SO) %>%
  z_score_count(HLD) %>%
  z_score_rate(ERA, IP, TRUE) %>%
  z_score_rate(WHIP, IP, TRUE) %>%
  mutate(
    z_sum = pmap_dbl(
      select(., starts_with("z_")), 
      sum
    ),
    type = "pitcher") %>% 
  select(Name, type, z_sum) -> proj_pitcher

```

#### Fantasy team rosters
```{r}
source("R/yahoo_api.R")
league_string <- paste0("388.l.",Sys.getenv("LEAGUE_KEY"))
get_team_roster(league_string, 8) %>% 
  clean_names() %>% 
  mutate(owner = 'Tom',
         name_full = chartr("áéóíñú", "aeoinu", name_full)) -> my_team
get_team_roster(league_string, 3) %>% 
  clean_names() %>% 
  mutate(owner = 'Chris',
         name_full = chartr("áéóíñú", "aeoinu", name_full)) -> chris_team

my_team %>% 
  bind_rows(chris_team) %>% 
  select(owner, name_full, display_position) -> roster_players

```

#### Draft results
```{r}
draft_results_raw <- read_tsv("data/draftResults.tsv") %>% 
  clean_names()

draft_results <- draft_results_raw %>% 
  select(pick, player)
```

#### Join data
```{r}
projections <- proj_batter %>% 
  bind_rows(proj_pitcher)

roster_players %>% 
  left_join(draft_results, by = c("name_full" = "player")) %>% 
  left_join(projections, by = c("name_full" = "Name")) -> trade_df

trade_df %>% saveRDS("data/tradeReview_20190405.Rds")
```