### 2019-04-09 rosterReview.Rmd
Laying ground work for roster review

---
title: "Roster Review"
author: "Tom Kain"
date: "April 5, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(glue)
library(plotly)
```

Various general roster reviews throughout the year.

Long and short term views for addressing roster gaps, potential trade targets, free agents, etc.