### 2019-04-09 rosterReview.Rmd
Moving often used code to R scripts, adding new projections

---
title: "Roster Review"
author: "Tom Kain"
date: "April 9, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(glue)
library(plotly)
source("R/z_scores.R")
source("R/yahoo_api.R")
```

Various general roster reviews throughout the year.

Long and short term views for addressing roster gaps, potential trade targets, free agents, etc.

```{r}
if (!file.exists("data/rosters_with_projections_20190409.Rds")) {

  get_team_rosters() -> rosters

  proj_batter <- add_z_scores_to_projections("data/depthChartProjections_ros_batters_20190409.csv","batter") 
  proj_pitcher <- add_z_scores_to_projections("data/depthChartProjections_ros_pitchers_20190409.csv", "pitcher")

  draft_results <- read_tsv("data/draftResults.tsv") %>% 
    clean_names() %>% 
    mutate(player = chartr("áéóíñú", "aeoinu", player)) %>% 
    select(pick, player)

  projections <- proj_batter %>% 
  bind_rows(proj_pitcher)

  projections$Name[projections$Name == 'Enrique Hernandez'] <- "Kike Hernandez"
  projections$Name[projections$Name == 'Peter Alonso'] <- "Pete Alonso"
  projections$Name[projections$Name == 'Joshua James'] <- "Josh James"

  rosters %>% 
    left_join(draft_results, by = c("name_full" = "player")) %>% 
    full_join(projections, by = c("name_full" = "Name")) %>% 
    mutate(team = if_else(is.na(team),"Free Agent",team),
           team = fct_reorder(team, z_sum)) -> rosters_with_projections

  rosters_with_projections %>% 
    saveRDS("data/rosters_with_projections_20190409.Rds")
} else {
  rosters_with_projections <- readRDS("data/rosters_with_projections_20190409.Rds")
}

```

```{r}
rosters_with_projections %>% 
  mutate(z_sum = round(z_sum)) %>% 
  ggplot(aes(x=team,y=z_sum,color=type)) +
  geom_jitter(aes(text=glue("{name_full} - {display_position}", color=type))) +
  coord_flip() -> p

ggplotly(p)

rosters_with_projections %>% 
  filter(!type == 'NA', !team == 'Free Agent') %>%
  ggplot(aes(x=type,y=z_sum)) +
  geom_boxplot()

rosters_with_projections %>% 
  filter(!type == 'NA', !team == 'Free Agent') %>%
  group_by(type) %>% 
  summarize(adj = median(z_sum)) -> z_sum_adj

rosters_with_adj_projections <- rosters_with_projections %>% 
  filter(!type == 'NA') %>% 
  mutate(zar = if_else(type=="batter",
                       z_sum-z_sum_adj$adj[z_sum_adj$type=="batter"],
                       z_sum-z_sum_adj$adj[z_sum_adj$type=="pitcher"]))

rosters_with_adj_projections %>% 
  ggplot(aes(x=type,y=zar)) +
  geom_boxplot()

rosters_with_adj_projections %>% 
  mutate(zar = round(zar)) %>% 
  ggplot(aes(x=team,y=zar,color=type)) +
  geom_jitter(aes(text=glue("{name_full} - {display_position}", color=type))) +
  coord_flip() -> z

ggplotly(z)

rosters_with_adj_projections %>% 
  filter((team == 'Free Agent' & zar >=0) | (grepl("Plouffe",team) & zar < 0)) %>% 
  arrange(zar %>% desc) %>% 
  select(name_full, display_position, type, zar) %>% 
  View()
```

```{r}
rosters_with_adj_projections %>% 
  filter(grepl('2B',display_position)) %>% 
  ggplot(aes(x=team,y=zar,color=team)) +
  geom_point(aes(text=name_full)) +
  coord_flip() -> sb

ggplotly(sb)

rosters_with_adj_projections %>% 
  filter(grepl('3B',display_position)) %>% 
  ggplot(aes(x=team,y=zar,color=team)) +
  geom_point(aes(text=name_full)) +
  coord_flip() -> tb

ggplotly(tb)

rosters_with_adj_projections %>% 
  mutate(third_base = grepl('3B',display_position),
         second_base = grepl('2B',display_position),
         position_plot_display = case_when(
           third_base & second_base ~ 'Both',
           third_base ~ '3B',
           second_base ~ '2B',
           TRUE ~ "Other")) %>% 
  filter(!position_plot_display == 'Other') %>% 
  mutate(team = fct_relevel(team, "Plouffe There It Is!")) %>% 
  ggplot(aes(x=team,y=zar,color=position_plot_display)) +
  geom_point(aes(text=name_full)) +
  coord_flip() -> both

ggplotly(both)

# free agents?

```