### 2019-07-04 midyear_perfVsProj.Rmd
Starting comparison of projections vs this year's performance

---
title: "Midyear Performance vs Projection"
author: "Tom Kain"
date: "7/3/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(glue)
library(plotly)

source("R/helper.R")
```

```{r}
#Latest rosters with projections
pitchers <- load_players_with_projections() %>% 
  filter(type == 'pitcher')

#Manually loading latest performance from fangraphs
pitcher_perf <- "temp/FanGraphs Leaderboard.csv" %>% 
  add_z_scores("pitcher") %>% 
  clean_names() %>% 
  select(type, playerid, name:so,era:war, z_sv:z_sum) %>% 
  mutate(playerid = as.character(playerid))

rosters_perf <- pitcher_perf %>% 
  rename(z_perf = z_sum) %>% 
  select(playerid, z_perf) %>% 
  right_join(pitchers)

rosters_perf %>% 
  filter(team != 'Free Agent') %>% 
  summarize(perf_adj = median(z_perf, na.rm = TRUE),
            proj_adj = median(z_sum, na.rm = TRUE)) %>% 
  summarize(adj = proj_adj - perf_adj) %>% 
  pull() -> adj

rosters_perf %>% 
  mutate(z_perf = z_perf + adj) -> rosters_perf
```

```{r}
rosters_perf %>% 
  filter(grepl("SP",positions)) %>% 
  ggplot(aes(z_sum,z_perf,color=team)) +
  geom_point(aes(text=name_full)) +
  geom_abline(slope = 1)-> p

ggplotly(p)
```