### 2021-03-03 Rmd/league_scoring_settings.Rmd
Evaluate league scoring settings by scraping 2020 performance data from baseball reference, plotting the distribution of points scored per player and position

---
title: "League Scoring Settings"
author: "Tom Kain"
date: "3/2/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
source(here::here("R", "scrape_bref.R"))
```

```{r import}

year <- 2020
batting_raw <- player_standard_batting(year)
pitching_raw <- player_standard_pitching(year)
starting_pitching_raw <- player_starting_pitching(year)
relief_pitching_raw <- player_relief_pitching(year)
```

```{r clean}

clean <- function(df) {
  df %>% 
    filter(name != 'Name') %>% 
    mutate(rk = as.numeric(rk)) %>% 
    group_by(name) %>% 
    arrange(name, rk) %>% 
    mutate(rnk = row_number()) %>% 
    filter(rnk == 1) %>% 
    ungroup()
}

batting <- 
  batting_raw %>% 
  clean() %>% 
  select(-rk, -rnk, -name, -age, -tm, -lg) %>% 
  sapply(as.numeric) %>% 
  as_tibble() %>% 
  filter(g >= 10)

starting_pitching <- 
  starting_pitching_raw %>% 
  clean() %>% 
  select(name, qs)

relief_pitching <-
  relief_pitching_raw %>% 
  clean() %>% 
  select(name, b_sv) %>% 
  rename(bsv = b_sv)

pitching <-
  pitching_raw %>% 
  clean() %>% 
  left_join(starting_pitching, by = "name") %>% 
  left_join(relief_pitching, by = "name") %>% 
  select(-name, -rk, -rnk, -age, -tm, -lg) %>% 
  sapply(as.numeric) %>% 
  as_tibble() %>% 
  filter(g >= 10) %>% 
  mutate(qs = if_else(is.na(qs), 0, qs),
         bsv = if_else(is.na(bsv), 0, bsv))

```

```{r scoring rules}

batting <- 
  batting %>% 
  mutate(x1b = h - x2b - x3b - hr,
         tb = x1b + (2 * x2b) + (3 * x3b) + (4 * hr),
         points = r + x1b + (2 * x2b) + (3 * x3b) + (4 * hr) + tb + (2 * rbi) + bb + (-3 * so) + (2 * sb) + (-2 * cs))

pitching <-
  pitching %>% 
  mutate(ip_whole = round(ip),
         points = (ip_whole * 3) + (-1 * h) + (-2 * er) + (-1 * bb) +  so + (5 * qs) + (10 * sho) + (10 * w) + (-5 * l) + (7 * sv) + (-3 * bsv))

```

```{r point results}

points <-
  batting %>% 
  group_by(points) %>% 
  summarize(batting = n()) %>% 
  ungroup() %>% 
  pivot_longer(!points, names_to = "position", values_to = "count")

points <- 
  pitching %>% 
  group_by(points) %>% 
  summarize(pitching = n()) %>% 
  ungroup() %>% 
  pivot_longer(!points, names_to = "position", values_to = "count") %>% 
  bind_rows(points)

points %>% 
  ggplot(aes(x = points, group=position, color=position)) +
  geom_freqpoly()

```